---
title: "Homework No 3"
author:
- UTRGV School of Mathematical and Statistical Sciences
- STAT 3301 Applied Statistics
- Samuel Lee
output:  html_notebook
---

---

>**Objective:** Use R studio to solve problems involving discrete random
variables and applications to Binomial and Poisson distributions.

---

#### Problem 1

Refer to the accompanying table, which describes results from groups of 8 births from 8
different sets of parents. The random variable $X$ represents the number of girls among 8 children. (Hint: you may use R function such as `sum`, `*` to do your computations)

```{r, echo=FALSE}
m <- c(0:8, 0.005, 0.031, 0.109, 0.219, 0.273, 0.219, 0.109, 0.031, 0.004)
m <- matrix(m, nrow = 2, byrow = T)
df <- as.data.frame(m,  row.names=c('x', 'P(x)'))
colnames(df) <- rep('', 9)
knitr::kable(df)
```

* Show that $P(X)$ is a valid probability discrete probability mass function. (Hint: there are two conditions that need to be satisfied For one of these criteria you may use function `rowSums`)

```{r}
cat("Has to be non-negative. We can just check which P(x) is greater than 0 \n")
cat(df[2,] > 0)
cat("\n Since they are all true no values are less than 0 which means all values are non-negative. \n")
cat("\n")
cat("P(x) has to sum to 1. Here is what out P(x) sums to: \n")
sum(df[2,])
```


* Display a graphical representation of the pmf.

```{r}
x <- 0:8
plot(x,dbinom(x ,size = 9,prob = 1/9),type="h", ylab="PMF", xlab="number of births", 
     main="Binomial Distribution PMF")
```

* Find the mean and standard deviation for the numbers of girls in 8 births.

```{r}
cat('Find the mean for the numbers of girls in 8 births: \n')
mean_8girls <- sum(df[1,]*df[2,])
print(mean_8girls)
cat('Find the standard deviation for the numbers of girls in 8 births: \n')
x_squared = df[1,]^2
#print(x_squared)
sqrt(sum(x_squared*df[2,])-(mean_8girls^2))

```

* Find the probability of getting 7 or more girls in 8 births.

```{r}
cat("The probability of getting 7 or more girls in 8 births: \n")
df[2,8]+df[2,9]
```

* Find the probability of getting 6 or more girls in 8 births.

```{r}
cat("The probability of getting 6 or more girls in 8 births: \n")
df[2,8]+df[2,9]+df[2,7]
```

* Find the probability of getting exactly 1 girl in 8 births.

```{r}
cat("Find the probability of getting exactly 1 girl in 8 births: \n")
df[2,2]
```

* If the value $x^*$  is greater than the 0.95 quantile, we say that $x^*$ is significantly high. Is 6 a significantly high number? Is 7 a significantly high number?

```{r}
cat("Neither 6 nor 7 are greater than the 0.95 quartile.")
cat("\n")
8*.95
cat("You have to be greater than 7.6, which is the number of entries times 0.95.")
```




#### Problem 2

Let $X$ be a Bernoulli random variable such that $x=0$ with probability $1-p,$ and 
$x=1$ with probability $p.$ In mathematical notation, this is the same as $X\sim \text{Bernoulli}(p).$

* Compute $E(X^2).$

```{r}
cat("E(X^2) = (1^2)*p + (0^2)*(1-p) = p")
```


* Show that $V(X) =  p(1- p).$ (Hint: is short formula for variance)

```{r}
cat("Var(X) = E(X^2) - [E(X)]^2 = p - p^2 = p(1-p)")
```


* Compute $E(X^{101}).$

```{r}
cat("E(X^101) = (1^101)*p + (0^101)*(1-p) = p")
```



#### Problem 3

The article “Should You Report That Fender-Bender?” (Consumer Reports, Sept. 2013: 15) reported that 7 in 10 auto accidents involve a single vehicle (the article recommended always reporting to the insurance company an accident involving multiple vehicles). Suppose 15 accidents are randomly selected. Use R functions to answer the following questions (Hint: Model this as a Bernoulli distribution):

* What is the probability that at most 4 involve a single vehicle?

```{r}
cat("What is the probability that at most 4 involve a single vehicle? \n")
pbinom(4,15,0.7)
```

* What is the probability that exactly 4 involve a single vehicle?

```{r}
cat("What is the probability that exactly 4 involve a single vehicle?\n")
dbinom(4,15,0.7)
```

* What is the probability that exactly 6 involve multiple vehicles?

```{r}
cat("What is the probability that exactly 6 involve multiple vehicles? \n")
dbinom(9,15,0.7)
```

* What is the probability that between 2 and 4, inclusive, involve a single vehicle?

```{r}
cat("What is the probability that between 2 and 4, inclusive, involve a single vehicle? \n")
asdf <- pbinom(4,15,0.7) - pbinom(1,15,0.7)
cat(asdf)
```

* What is the probability that at least 2 involve a single vehicle?

```{r}
cat("What is the probability that at least 2 involve a single vehicle? \n")
1-pbinom(1,15,0.7)
```



#### Problem 4

53% of adults use eyeglasses for vision correction (based on data
from a Vision Council survey). Four adults are randomly selected.

* Use the multiplication rule to find the probability that the first three use
eyeglasses and the fourth does not use eyeglasses. That is, find P(EEEN), where E
denotes an adult who uses eyeglasses and N denotes an adult who does not use
eyeglasses. (Hint: Trials in Binomial experiments are independent)

```{r}
cat("Probability of P(EEEN)\n")
gamerenergy <- .53*.53*.53*.47
cat(gamerenergy)
```


* Beginning with EEEN, make a complete list of the different possible arrangements
of three adults who use eyeglasses and one who does not use eyeglasses; then find
the probability for each entry in the list.

```{r}
cat("Probability of P(EEEN)\n")
cat(gamerenergy)
cat("\nProbability of P(EENE)\n")
cat(gamerenergy)
cat("\nProbability of P(ENEE)\n")
cat(gamerenergy)
cat("\nProbability of P(NEEE)\n")
cat(gamerenergy)
```


* Based on the preceding results, what is the probability of getting exactly three
adults who use eyeglasses and one who does not? (Hint: Binomial outcomes are mutually exclusive)

```{r}
cat("exactly three adults and one who does not\n")
cat(gamerenergy)
```



#### Problem 5

In a recent year, there were 4229 births at NYU Langone Medical Center (based on
data from the NYU Langone website). Assume that the number of births each day is
about the same, and assume that the Poisson distribution is a suitable model.

* Find $\mu,$ the mean number of births per day.

```{r}
cat("mu\n")
mu <- 4229/365
cat(mu)
```


* Find the probability that on a randomly selected day, there are exactly 10
births. That is, find $P(10)$, where $P(x)$ is the probability of $x$ births in a day.

```{r}
cat("P(10)\n")
p_10 = (mu^10)*(exp(-mu))/(factorial(10))
cat(p_10)
```


  * Use the Poisson distribution formula
  
```{r}
cat("P(10) using the Poisson distribution formula\n")
p_10 = (mu^10)*(exp(-mu))/(factorial(10))
cat(p_10)
```
  
  * Use the an R function for this calculation.
  
```{r}
cat("P(10) using R function: \n")
dpois(x=10,lambda=mu)
```
  
  

#### Problem 6

During the past 34 years, there were 56 cases of diphtheria in the United
States. (Hint: Use the Poisson distribution)

* Find the mean number of cases of diphtheria per year. Express the result with
five decimal places.

```{r}
mew = 56/34
cat("This is how many diphteria cases per year: \n")
sprintf(mew, fmt = '%#.5f')
```


* Find the probability of no cases of diphtheria in a year.

```{r}
cat("Probability of no cases of diphtheria: \n")
dpois(x=0,lambda=mew)
```

* Find the probability that the number of diphtheria cases in a year is 5 or
fewer. If a year has more than 5 cases of diphtheria, is that a significantly high
number?

```{r}
cat("Probability that number of cases is 5 or fewer: \n")
dpois(x=0,lambda=mew)+dpois(x=1,lambda=mew)+dpois(x=2,lambda=mew)+dpois(x=3,lambda=mew)+dpois(x=4,lambda=mew)+dpois(x=5,lambda=mew)
cat("\n If a year has more than 5 cases, is that a significantly high number? \n ")
cat("\nYes because having 5 or fewer cases makes up 99.3%, meaning you only have 0.7% chance of getting more than 5 cases.")
```








