---
title: "Homework No 2"
author:
- UTRGV School of Mathematical and Statistical Sciences
- STAT 3301 Applied Statistics
- Samuel Lee
output:  html_notebook
---

---

>**Objective:** Use the R notebook to present and describe data using descriptive statistics techniques.
In particular, you will use the built in dataset `iris` for descriptive statistics.

---


* Use the R studio built in dataset `iris` for this homework. For example, use the command `df <- iris` to
create a data frame called `df` which now contains the records of the iris data set.

```{r}
#complete
df <- iris
```


* Use functions such as `dim` and answer how many data observations do this data set have? How many variables, and what type of variables they are (e.g., qualitative, etc)?'

There are 5 type of variables. "Species" is the name of the plant, and is therefore qualitative. The rest are quantitative.

```{r}
#complete
dim(df)
str(df)
```


* What is the value of the Petal length for observation 42?

According to the following code, Petal length for observation 42 is 1.3 (since R starts at 1, not 0).

```{r}
#complete
petal_lengths <- df$Petal.Length
petal_lengths[42]
```


* What is the Species for observation No 51?
```{r}
#complete
species <- df$Species
species[51]
#according to this, species is versicolor
```


* Use the `head` and `summary` functions on `df` to have feeling of the data in front of you. Comment
on your findings.

```{r}
#complete
#according to this, Setosa, Versicolor, and Virginia are our three species. It appears we have data for 50 of each species.
head(df)
summary(df)
```


* Use diagrams of stem and leaves,  histogram and boxplot to investigate the distribution of the variable
`Sepal.Width` across all Species, that is, for all observations. What are your comments?

```{r}
#complete
width <- df$Sepal.Width

#According to the stem plot, most plants are within 28 and 32
stem(width)

#According to the histogram, the most frequent widths of these plants is just under 3.0
hist(width)

#According to the boxplot, the Median is somewhere around 3.0, the upper quartile is around 3.3 and the lower quartile is around 2.7
boxplot(width)
```


* Compute the mean of `Sepal.Width` for Species Setosa, Versicolor and Virginica. That is, you need to compute the mean (use the function `mean`) three times, one for Setora, one for Versicolor and one for Virginical. This also mean that you need to slice or index your data.

```{r}
#complete
library(rlang)
library(dplyr)
output <- iris %>% 
         group_by(Species) %>% 
         summarize_each(funs(mean, sd))
`colnames<-`(t(output[3]), as.character(output$Species))
```


* Use the `boxplot` function for all three species. You will have to see three boxplots in one plot. Are the mean you computed in previous point consistent?

```{r}
boxplot(output$Petal.Width_mean[1], output$Petal.Width_mean[2], output$Petal.Width_mean[3])

#There appears to be an outlier.
```


* There seems to be an outlier for Sepal Width in setosa species. Plot a histogram of `Sepal.Widht` only
for Setosa species. What is the exact value for this outlier? 

```{r}
#complete
Temp = head(df$Sepal.Width , 50)
hist(Temp)

#According to summary, the outlier is a width of 4.400
summary(Temp)
```



