---
title: "Homework No 1"
author:
- UTRGV School of Mathematical and Statistical Sciences
- STAT 3301 Applied Statistics
- (Samuel Lee)
output:  html_notebook
---

---


>**Objective:** Use the R notebook to present homework reports in a standard and organized manner.


- Read Chapters No 1 and No 2 of Book:	R Cookbook, 2nd Edition by Paul Teetor ISBN-10: 1492040681 https://rc2e.com/

- Review the documentation for the function `seq`. Use the `seq` function in R to create a numeric
vector with 150 values ranging from $0$ to $2\pi.$ Hint: check the parameter `length.out` in the
`seq` function. Assign your result to a variable named `t`.

- Compute the sine of `t` using the function `sin`. Assign your result to variable `fs.`

- What is the proportion of values in `fs` that are strictly greater than 0?

- What is the mean of `fs`?

- Create a vector named `fs2` that corresponds with the square of `fs`. 

- Create a Boolean vector named `idx` with entries `TRUE` where  `fs2>0.3`, an FALSE otherwise.

- What is the mean of `fs` considering only entries with `fs2>0.3`? Hint: use `idx`

- Use `plot` function to plot `fs` vs `t`

- Improve the graphics according to your taste. Eg., look the documentation for `plot`, use parameters like `type`, `col`, `lwd`, etc. Include title (using `main`), `xlab` and `ylab`.

---
Start your homework here:


this is how we create t:
```{r}
t <- seq(1, 2 * pi, length.out = 150)
print(t)
```


this is how we create fs:
```{r}
fs <- sin(t)
print(fs)
```
```{r}
length(fs[fs>0])
61/150
```

According to length function, fs has 61 elements that are greater than 0. So 40.67% of elements are greater than 0.
```{r}
mean(fs)
```
According to mean function, -0.08361743 is the mean of fs.

This is how we create fs2:
```{r}
fs2 <- fs ^ 2
print(fs2)
```

This is how we create idx:
```{r}
idx <- fs2 > 0.3
print(idx)
```

Mean of fs using only fs2>.3 is:
```{r}
mean <- 0
for (x in idx){
for (y in fs){
    if (x==TRUE)
  {
    mean <- mean + y
  }
}
}
print(mean)
```


plot of fs vs t:
```{r}
plot(fs,t, main = "sin(t) vs t", xlab = "sin(t)", ylab = "t", type = "l")
```

